---
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# vachette
A method for visualization of PMx models.

Vachette builds off previous work done in [V2ACHER](https://ascpt.onlinelibrary.wiley.com/doi/10.1002/psp4.12679), providing a method to visualize PKPD analyses which are impacted by covariate effects.

## Installation

**Installation instructions will be available once the Github repository has been made public**

## Examples

### IV

```{r iv}
# Import Example Data
obs.data <- read.csv(system.file(package = "vachette", "examples", "iv-obs.csv")) |> 
  dplyr::rename(WT = "vachette.cov1")
typ.data <- read.csv(system.file(package = "vachette", "examples", "iv-typ.csv")) |> 
  dplyr::rename(WT = "vachette.cov1")

library(vachette)

vd <-
  vachette_data(
    obs.data,
    typ.data,
    covariates = c("WT" = 70),
    ref.dosenr = 1,
    iiv.correction = FALSE,
    model.name = "Intravenous (i.v)"
  )

vd <- vd |>
  apply_transformations()

p.vachette(vd)
  
```

### Oral Two-Cov 

```{r oral_two_cov}
obs.data <- read.csv(system.file(package = "vachette", "examples", "oral-two-cov-obs.csv"))
typ.data  <- read.csv(system.file(package = "vachette", "examples", "oral-two-cov-typ.csv"))

vd <-
  vachette_data(
    obs.data,
    typ.data,
    covariates =  c("WT" = 70, "AGE" = 30),
    ref.dosenr = 1,
    iiv.correction = FALSE,
    model.name = "oral-two-cov"
  )

vd <- vd |>
  apply_transformations(window = 23,
                        window.d1.refine = 7,
                        window.d2.refine = 5)

p.vachette(vd)

```

### Sigmoid

```{r sigmoid}
obs.data <- read.csv(system.file(package = "vachette", "examples", "sigmoid-obs.csv")) |>
  dplyr::rename("ID" = id)
typ.data  <- read.csv(system.file(package = "vachette", "examples", "sigmoid-typ.csv"))

vd <-
  vachette_data(
    obs.data,
    typ.data,
    covariates =  c("vachette.cov1" = 70),
    ref.dosenr = 1,
    iiv.correction = FALSE,
    model.name = "sigmoid"
  )

vd <- vd |>
  apply_transformations(window = 17,
                        window.d1.refine = 7,
                        window.d2.refine = 5)

p.vachette(vd)


```


### Oral-Absorption

```{r oral_absorption}
obs.data <- read.csv(system.file(package = "vachette", "examples", "oral-absorption-obs.csv"))
typ.data  <- read.csv(system.file(package = "vachette", "examples", "oral-absorption-typ.csv"))

vd <-
  vachette_data(
    obs.data,
    typ.data,
    covariates =  c("vachette.cov1" = 70),
    ref.dosenr = 1,
    iiv.correction = FALSE,
    model.name = "oral-absorption"
  )

vd <- vd |>
  apply_transformations(window = 17,
                        window.d1.refine = 7,
                        window.d2.refine = 5)

p.vachette(vd)

```


### Pembro

**Note: The Pembro data set contains a categorical covariate e.g., `SCHED`**

```{r pembro, message=FALSE}
obs.data <- read.csv(system.file(package = "vachette", "examples", "pembro-obs.csv"))
typ.data  <- read.csv(system.file(package = "vachette", "examples", "pembro-typ.csv"))

vd <-
  vachette_data(
    obs.data,
    typ.data,
    covariates =  c(SCHED = 'Q3W', ALB = 53.5),
    ref.dosenr = 3,
    iiv.correction = FALSE,
    model.name = "pembro"
  )

print(vd)

```

If no reference value is specified, the covariate central tendency is used e.g., median for continuous and mode for categorical covariate.

```{r pembro2, message=FALSE}
vd <-
  vachette_data(
    obs.data,
    typ.data,
    covariates =  c("SCHED", "ALB"),
    ref.dosenr = 3,
    iiv.correction = FALSE,
    model.name = "pembro"
  )

print(vd)

```

```{r pembro3, message = FALSE}
vd <- vd |>
  apply_transformations(window = 17,
                        window.d1.refine = 7,
                        window.d2.refine = 5)

p.vachette(vd)
```
